---
title: "Final project analysis"
author: "Disha Bhaiya"
date: "3/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 
summary(Lead)

````

#Read file
```{r cars}
rm(list=ls())
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 
#Lead <- as.data.frame(scale(Lead))
set.seed(2018)
ss <- sample(1:3,size=nrow(Lead),replace=TRUE,prob=c(0.6,0.2,0.2))
train <- Lead[ss==1,-c(1)]
crossval <- Lead[ss==2,-c(1)]
test <- Lead[ss==3,-c(1)]

```

#Running linear regression on Full mode

```{r}



lin.fit = lm(MeanAQ_3days_after ~ . ,data = train)
summary(lin.fit)$r.sq
summary(lin.fit)

trainingpred = predict(lin.fit, train)

residual = trainingpred - train[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_train = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_train
NRMSE = RMSE_train/mean(train$MeanAQ_3days_after)
NRMSE2 = RMSE_train/(max(train$MeanAQ_3days_after)-min(train$MeanAQ_3days_after))

cvpred = predict(lin.fit, crossval)
residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_cv = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_cv

testpred = predict(lin.fit, test)
residual = testpred - test[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_test = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_test
```



# Training/Val/Test accuracy on full model
```{r}
setsize = seq(from = 100000, to = nrow(Lead), by = 5000)
#setsize
k = 1
Errortrain = rep(0,length(setsize))
Errortest = rep(0,length(setsize))
Errortcv = rep(0,length(setsize))
for (i in setsize) {
  
  trainingsetsize = i
  Leadt = sample(1:nrow(Lead), trainingsetsize)
  Leadtemp = Lead[Leadt,]
  ss <- sample(1:3,size=nrow(Leadtemp),replace=TRUE,prob=c(0.6,0.2,0.2))
  train <- Leadtemp[ss==1,]
  crossval <- Leadtemp[ss==2,]
  test <- Leadtemp[ss==3,]

  lin.fit = lm(MeanAQ_3days_after ~ . ,data = train)
  trainingpred = predict(lin.fit, train)
  residual = trainingpred - train[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortrain[k] = RMSE
  
  cvpred = predict(lin.fit, crossval)
  residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortcv[k] = RMSE
  
  testpred = predict(lin.fit, test)
  residual = testpred - test[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortest[k] = RMSE
  
  k = k+1
}

plot( setsize, Errortrain, type="l", col="red" , ylim = c(0.015,0.02), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortcv, type="l", col="green",ylim = c(0.015,0.02), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortest, type="l", col="blue",ylim = c(0.015,0.02), ylab = "Error")


````

## Log transform for fire size

#Running linear regression on Full mode

```{r}


lin.fit = lm(MeanAQ_3days_after ~  log(FIRE_SIZE) + Mean.1 + Mean.2 + Mean.3 + Mean.4 + Mean.5 + Mean...Day.of.Event + FIRE_YEAR + DISCOVERY_DOY + CONT_DOY + ddist , data = train)


trainingpred = predict(lin.fit, train)

residual = trainingpred - train[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_train = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_train

cvpred = predict(lin.fit, crossval)
residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_cv = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_cv

testpred = predict(lin.fit, test)
residual = testpred - test[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
RMSE_test = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_test
```



# Training/Val/Test accuracy on log model
```{r}
rm(list=ls())
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 

setsize = seq(from = 100000, to = nrow(Lead), by = 5000)
#setsize
k = 1
Errortrain = rep(0,length(setsize))
Errortest = rep(0,length(setsize))
Errortcv = rep(0,length(setsize))
for (i in setsize) {
  
  trainingsetsize = i
  Leadt = sample(1:nrow(Lead), trainingsetsize)
  Leadtemp = Lead[Leadt,]
  ss <- sample(1:3,size=nrow(Leadtemp),replace=TRUE,prob=c(0.7,0.15,0.15))
  train <- Leadtemp[ss==1,]
  crossval <- Leadtemp[ss==2,]
  test <- Leadtemp[ss==3,]
  lin.fit = lm(MeanAQ_3days_after ~  log(FIRE_SIZE) + Mean.1 + Mean.2 + Mean.3 + Mean.4 + Mean.5 + Mean...Day.of.Event + FIRE_YEAR + DISCOVERY_DOY + CONT_DOY + ddist ,data = train)
  trainingpred = predict(lin.fit, train)
  residual = trainingpred - train[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortrain[k] = RMSE
  
  cvpred = predict(lin.fit, crossval)
  residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortcv[k] = RMSE
  
  testpred = predict(lin.fit, test)
  residual = testpred - test[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortest[k] = RMSE
  
  k = k+1
}


plot( setsize, Errortrain, type="l", col="red" , ylim = c(0,0.0003), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortcv, type="l", col="green",ylim = c(0,0.0003), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortest, type="l", col="blue",ylim = c(0,0.0003), ylab = "Error")


```

#Subset Selection


```{r}
library(leaps)
library(MASS)


# Fit the full model 
lin.fit = lm(MeanAQ_3days_after ~ . ,data = train)
# Stepwise regression model
step.model <- stepAIC(lin.fit, direction = "both", 
                      trace = FALSE)
summary(step.model)
library(caret)
# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 12)
# Train the model
step.model2 <- train(MeanAQ_3days_after ~., data = train,
                    method = "leapForward", 
                    tuneGrid = data.frame(nvmax = 1:12),
                    trControl = train.control
                    )

```


# ### TRAINING AND TEST ERROR on limited model
```{r}



#lin.fit = lm(MeanAQ_3days_after ~ ddist + Mean.4 + Mean.3 +log(FIRE_SIZE) + Mean...Day.of.Event + DISCOVERY_DOY + CONT_DOY + FIRE_YEAR,data = train)

lin.fit = lm(MeanAQ_3days_after ~   Mean.1 + Mean.2 + Mean.4 + Mean.3 + Mean...Day.of.Event + DISCOVERY_DOY  + CONT_DOY + ddist + FIRE_YEAR,data = train)
trainingpred = predict(lin.fit, train) 
residual = trainingpred - train[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortrain = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortrain

cvpred = predict(lin.fit, crossval)
residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortcv = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortcv 

testpred = predict(lin.fit, test)
residual = testpred - test[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortest = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortest 

```

# Training/Val/Test accuracy on limited model
```{r}
rm(list=ls())
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 

setsize = seq(from = 100000, to = nrow(Lead), by = 5000)
#setsize
k = 1
Errortrain = rep(0,length(setsize))
Errortest = rep(0,length(setsize))
Errortcv = rep(0,length(setsize))
for (i in setsize) {
  trainingsetsize = i
  Leadt = sample(1:nrow(Lead), trainingsetsize)
  Leadtemp = Lead[Leadt,]
  ss <- sample(1:3,size=nrow(Leadtemp),replace=TRUE,prob=c(0.7,0.15,0.15))
  train <- Leadtemp[ss==1,c(2,4,5,7,8,9,10,11,13)]
  crossval <- Leadtemp[ss==2,c(2,4,5,7,8,9,10,11,13)]
  test <- Leadtemp[ss==3,c(2,4,5,7,8,9,10,11,13)]
  
  lin.fit = lm(MeanAQ_3days_after ~  Mean.4 + Mean.1 + Mean.3 +log(FIRE_SIZE) + Mean...Day.of.Event + DISCOVERY_DOY + CONT_DOY + FIRE_YEAR,data = train)
  trainingpred = predict(lin.fit, train)
  residual = trainingpred - train[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortrain[k] = RMSE
  
  cvpred = predict(lin.fit, crossval)
  residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortcv[k] = RMSE
  
  testpred = predict(lin.fit, test)
  residual = testpred - test[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortest[k] = RMSE
  
  k = k+1
}

plot( setsize, Errortrain, type="l", col="red" , ylim = c(0,0.0003), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortcv, type="l", col="green", ylim = c(0,0.0003), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortest, type="l", col="blue", ylim = c(0,0.0003), ylab = "Error")


```




#Polynomial

```{r}


lin.fit = lm(MeanAQ_3days_after ~ I(Mean.4^2) + I(Mean.1^2) + I(Mean.3^2) +log(FIRE_SIZE) + I(Mean.2^2)+ I(Mean...Day.of.Event^2) + I(DISCOVERY_DOY^(2)) + I(CONT_DOY^(2)) + I(FIRE_YEAR^(2)),data = train)
#summary(lin.fit)
trainingpred = predict(lin.fit, train)
residual = trainingpred - train[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortrain = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortrain

cvpred = predict(lin.fit, crossval)
residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortcv = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortcv 

testpred = predict(lin.fit, test)
residual = testpred - test[,c("MeanAQ_3days_after")] 
residual_sqr = residual^2
Errortest = sqrt(sum(residual_sqr)/length(residual_sqr)) 
Errortest  

```


#Accuracy for polynomial model


# Training/Val/Test accuracy on log model
```{r}
rm(list=ls())
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 

setsize = seq(from = 100000, to = nrow(Lead), by = 5000)
#setsize
k = 1
Errortrain = rep(0,length(setsize))
Errortest = rep(0,length(setsize))
Errortcv = rep(0,length(setsize))
for (i in setsize) {
  
ss <- sample(1:3,size=nrow(Lead),replace=TRUE,prob=c(0.7,0.15,0.15))
trainingsetsize = i
  Leadt = sample(1:nrow(Lead), trainingsetsize)
  Leadtemp = Lead[Leadt,]
train <- Lead[ss==1,c(2,4,5,7,8,9,10,11,13)]
crossval <- Lead[ss==2,c(2,4,5,7,8,9,10,11,13)]
test <- Lead[ss==3,c(2,4,5,7,8,9,10,11,13)]


lin.fit = lm(MeanAQ_3days_after ~ I(Mean.4^2) + I(Mean.1^1) + I(Mean.3^1) +log(FIRE_SIZE) + Mean...Day.of.Event + I(DISCOVERY_DOY^(1/2)) + I(CONT_DOY^(1/2)) + I(FIRE_YEAR^(1/2)),data = train)
  trainingpred = predict(lin.fit, train)
  residual = trainingpred - train[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortrain[k] = RMSE
  
  cvpred = predict(lin.fit, crossval)
  residual = cvpred - crossval[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortcv[k] = RMSE
  
  testpred = predict(lin.fit, test)
  residual = testpred - test[,c("MeanAQ_3days_after")] 
  residual_sqr = residual^2
  RMSE = sqrt(sum(residual_sqr)/length(residual_sqr)) 
  Errortest[k] = RMSE
  
  k = k+1
}

plot( setsize, Errortrain, type="l", col="red" , ylim = c(0,0.0003),ylab = "Error")
par(new=TRUE)
plot( setsize, Errortcv, type="l", col="green",ylim = c(0,0.0003), ylab = "Error")
par(new=TRUE)
plot( setsize, Errortest, type="l", col="blue",ylim = c(0,0.0003), ylab = "Error")


```


##Neural Network

```{r}
# Read the Data
library(neuralnet)
trainNN = train
crossvalNN = crossval
testNN = test

NN = neuralnet(MeanAQ_3days_after ~ Mean.4 + Mean.2 + Mean.1 + Mean.3 +FIRE_SIZE + Mean...Day.of.Event + DISCOVERY_DOY + CONT_DOY + FIRE_YEAR, trainNN, hidden = 2 ,stepmax = 1e+08 , threshold = 0.01)

#plot neural network
plot(NN)

# Calculate Root Mean Square Error (RMSE)
predict_trainNN = compute(NN, trainNN)
Train.RMSE.NN = (sum((trainNN[,c(12)] - predict_trainNN$net.result)^2) / nrow(trainNN)) ^ 0.5
Train.RMSE.NN

predict_crossvalNN = compute(NN, crossvalNN)
Crossval.RMSE.NN = (sum((crossvalNN[,c(12)] - predict_crossvalNN$net.result)^2) / nrow(crossvalNN)) ^ 0.5
Crossval.RMSE.NN

predict_testNN = compute(NN, testNN)
Test.RMSE.NN = (sum((testNN[,c(12)] - predict_testNN$net.result)^2) / nrow(testNN)) ^ 0.5
Test.RMSE.NN

```


#Lambda

```{r}

library(glmnet)

lambdas <- 10^seq(-2, 3, by = .1)

x = as.matrix(train[,-c(12)])
y = train[,c(12)]

fit <- glmnet(x, y, alpha = 0, lambda = lambdas)
cv_fit <- cv.glmnet(x, y, alpha = 0, lambda = lambdas)
plot(cv_fit)
opt_lambda <- cv_fit$lambda.min
opt_lambda


y_predicted <- predict(fit, s = opt_lambda, newx = x)
residual = y_predicted - y 
residual_sqr = residual^2
RMSE_train = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_train

xcv = as.matrix(crossval[,-c(12)])
ycv = crossval[,c(12)]
y_predicted <- predict(fit, s = opt_lambda, newx = xcv)
residual = y_predicted - ycv 
residual_sqr = residual^2
RMSE_cv = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_cv

xt = as.matrix(test[,-c(12)])
yt = test[,c(12)]
y_predicted <- predict(fit, s = opt_lambda, newx = xt)
residual = y_predicted - yt 
residual_sqr = residual^2
RMSE_test = sqrt(sum(residual_sqr)/length(residual_sqr)) 
RMSE_test


```




******************************

#####
```{r}
rm(list=ls())
Lead = read.csv("Final_dataset_US_Lead.csv")
Lead = Lead[,(which(colSums(Lead) != 0))] 
Lead <- as.data.frame(scale(Lead))
k = 1
setsize = seq(from = 100000, to = nrow(Lead), by = 10000)
Error = rep(0,length(setsize))
Train.RMSE.NN= rep(0,length(setsize))
Crossval.RMSE.NN= rep(0,length(setsize))
Test.RMSE.NN= rep(0,length(setsize))


for (i in setsize) {
        trainingsetsize = i
        Leadt = sample(1:nrow(Lead), trainingsetsize)
        Leadtemp = Lead[Leadt,]
        t1 = sample(1:nrow(Leadtemp), 0.6*nrow(Leadtemp))
        t2 = setdiff(1:nrow(Leadtemp),t1)
        excepttrain = Leadtemp[t2,]
        t3 = sample(1:nrow(excepttrain), 0.5*nrow(excepttrain))
        t4 = setdiff(1:nrow(excepttrain), t3)
        trainNN = Leadtemp[t1,-c(1)]
        crossvalNN = Leadtemp[t3,-c(1)]
        testNN = Leadtemp[t4,-c(1)]
        rm(excepttrain,t1, t2, t3, t4,Leadtemp)
        # fit neural network
        set.seed(2)
        NN = neuralnet(MeanAQ_3days_after ~ ., trainNN, hidden = 3 ,stepmax = 1e+05 , threshold = 0.5)
        
        # plot neural network
        #plot(NN)
        
        # Calculate Root Mean Square Error (RMSE)
        predict_trainNN = compute(NN, trainNN[,-c(12)])
        Train.RMSE.NN[k] = (sum((trainNN[,c(12)] - predict_trainNN$net.result)^2) / nrow(testNN)) ^ 0.5
        #Train.RMSE.NN
        
        predict_crossvalNN = compute(NN, crossvalNN[,-c(12)])
        Crossval.RMSE.NN[k] = (sum((crossvalNN[,c(12)] - predict_crossvalNN$net.result)^2)/nrow(testNN))^ 0.5
        #Crossval.RMSE.NN
        
        predict_testNN = compute(NN, testNN[,-c(12)])
        Test.RMSE.NN[k] = (sum((testNN[,c(12)] - predict_testNN$net.result)^2) / nrow(testNN)) ^ 0.5
        #Test.RMSE.NN
         k=k+1
    
}



plot( setsize, Train.RMSE.NN, type="l", col="red" ,ylab = "Error")
par(new=TRUE)
plot( setsize, Crossval.RMSE.NN, type="l", col="green",ylab = "Error")
par(new=TRUE)
plot( setsize, Test.RMSE.NN, type="l", col="blue",ylab = "Error")

```
